---
title: "pres#2"
output: html_document
date: "2026-02-20"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
setwd("/Users/ppower1/Desktop/")
```

## Setup

```{r}
library("readr")
library("dplyr")
library("ggplot2")
library("fixest")
library(tidyr)
```

(b)
```{r cars}
df <- read_csv("problem_set_two_data.csv")
head(df)
```
(c)
```{r}
df <- df %>% filter(MA > 0)
```

(d)
```{r}
df <- df %>%
  group_by(fips) %>%
  filter(n_distinct(year) == 3) %>%
  ungroup()
```

## Part One

(a)
```{r}
df$log_revenue <- log(df$production_value)
df$log_MA <- log(df$MA)
```

(b)
```{r}
reg <- feols(log_revenue ~ log_MA | fips + state^year, data = df)
summary(reg)
```

(c)
```{r}
sd_change <- df %>%
  filter(year %in% c(1860, 1880)) %>%
  select(fips, year, log_MA) %>%
  pivot_wider(names_from = year, values_from = log_MA, names_prefix = "y") %>%
  mutate(delta = y1880 - y1860) %>%
  pull(delta) %>%
  sd()
sd_change
```

(d)
```{r}
beta_hat <- coef(reg)["log_MA"]
effect <- beta_hat * sd_change
effect
```

## Part Two

(a)
```{r}
df$capital_cost <- 0.10 * df$capital
df$log_K <- log(df$capital_cost)
df$log_L <- log(df$labor_cost)
df$log_M <- log(df$material_cost)
```

(b)
```{r}
alpha_K <- mean(df$capital_cost / df$production_value)
alpha_L <- mean(df$labor_cost / df$production_value)
alpha_M <- mean(df$material_cost / df$production_value)

alpha_K
alpha_L
alpha_M
```

(c)
```{r}
df$log_TFPR <- df$log_revenue - alpha_K * df$log_K - alpha_L * df$log_L - alpha_M * df$log_M
```

(d)
```{r}
reg_tfpr <- feols(log_TFPR ~ log_MA | fips + state^year, data = df)
summary(reg_tfpr)
```

```{r}
beta_tfpr <- coef(reg_tfpr)["log_MA"]
effect_tfpr <- beta_tfpr * sd_change
effect_tfpr
```


## Part Three
(a)
```{r}
# Compute sd by year
sd_by_year <- df %>%
  group_by(year) %>%
  summarize(sd_log_TFPR = sd(log_TFPR))

# Bar chart
ggplot(sd_by_year, aes(x = factor(year), y = sd_log_TFPR)) +
  geom_col() +
  labs(x = "Year", y = "SD of log(TFPR)",
       title = "Dispersion of Revenue Productivity Over Time") +
  theme_minimal()
```

(b)
The dispersion decreases by roughly 11% over thie time period
```{r}
sd_1860 <- sd_by_year$sd_log_TFPR[sd_by_year$year == 1860]
sd_1880 <- sd_by_year$sd_log_TFPR[sd_by_year$year == 1880]
pct_change <- (sd_1880 - sd_1860) / sd_1860 * 100
pct_change
```

(c)
Bar chart suggests that factors of production became more efficiently allocated over time. Our previous results suggest that railroads played a significant role in this regrard. They contributed to growth in the U.S. by improving market access which allowed for a more efficient allocation of factors of production.
